STAT0017 ICA 1 2018-19

### 2019-03-22 STAT0017_ICA1/SN17052480.Rmd
plot pot
finished ICA1 selected topics

<!-- saved from url=(0100)https://moodle-1819.ucl.ac.uk/pluginfile.php/1342515/mod_resource/content/7/stat0017ica1template.Rmd -->
<html><head><meta http-equiv="Content-Type" content="text/html; charset=UTF-8"></head><body>---
title: "STAT0017 ICA 1 2018-19"
author: "Student number: xxxxxxxx"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_width: 6
    fig_height: 4
---

```{r, include = FALSE}
# Load the data (you will need to edit this to point to the correct directory
# on your computer)
setwd("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1")
raw_wm <- load("wm.Rdata")
raw_pot <- load("pot.Rdata")
# You will also need to load any R libraries that you want to use.
library("evd")
library("evdbayes")
library("ismev")
library("revdbayes")
library("threshr")
library("bayesplot")
library("coda")
source("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1/pjnfunctions.R")
# Some parts of the analyses are based on simulation.
# We set a set a random number seed so that
# 1. The numeric results on which you comments will not change each time you run
#    your code (unless you change the order of the analyses)
# 2. Your work is easier for me to mark.
#### PLEASE DO NOT CHANGE THE FOLLOWING LINE OF CODE (KEEP THE SEED as 17) ####
set.seed(17)
```

@@

@@

```{r}
# Add R code here (and similarly elsewhere)
```

## Storm peaks (pot)

## Comments

@@

## GEV modelling of winter maxima

### Maximum Likelihood-Based Inference

@@

### Bayesian Inference

@@

## Binomial-GP modelling of storm peaks

### Threshold selection

### Maximum Likelihood-Based Inference

@@

### Bayesian Inference

@@

@@

## Build a GEV regression model
test123
${}$\hspace*{\fill} [15]

>>>>

<!-- saved from url=(0100)https://moodle-1819.ucl.ac.uk/pluginfile.php/1342515/mod_resource/content/7/stat0017ica1template.Rmd -->
<html><head><meta http-equiv="Content-Type" content="text/html; charset=UTF-8"></head><body>---
title: "STAT0017 ICA 1 2018-19"
author: "Student number: 17052480"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_width: 6
    fig_height: 4
---

```{r, include = FALSE}
# Load the data (you will need to edit this to point to the correct directory
# on your computer)
setwd("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1")
load("wm.Rdata")
load("pot.Rdata")
# You will also need to load any R libraries that you want to use.
library("evd")
library("evdbayes")
library("ismev")
library("revdbayes")
library("threshr")
library("bayesplot")
library("coda")
source("/Users/hongwei/Documents/GitHub/STAT/STAT0017_ICA1/pjnfunctions.R")
# Some parts of the analyses are based on simulation.
# We set a set a random number seed so that
# 1. The numeric results on which you comments will not change each time you run
#    your code (unless you change the order of the analyses)
# 2. Your work is easier for me to mark.
#### PLEASE DO NOT CHANGE THE FOLLOWING LINE OF CODE (KEEP THE SEED as 17) ####
set.seed(17)
```

@@

## Question 1 (a)
From the question we can see:
$$ F(x)= P(X \leq x)= \begin{cases} 1-e^{1/x}, & \text {for $x < 0$ .} \\ 1, & \text{for $x \geq 0$ .} \end{cases} \tag {*}$$
$$ M_n=max \lbrace x_1,...,x_n \rbrace $$ 
Hazard Function:
\begin{align}
h(x) & =\frac{1-F(x)}{f(x)} \\
  & = \frac{e^{1/x}}{e^{1/x} \times (-x^{-2})} \\
  & =-x^{2}
\end{align} 
Derivative of h(x):
$h^{'}= -2x$

\begin{align}
\lim_{x \to x^F} h^{'}(x) \to \xi \\
Let\ \xi = -2M_n \\
We\ can\ see:\  x \sim GEV(0,1,-2M_n)
\end{align}

From the slides we have:
$$\begin{cases} 1-F(b_n)=1/n \\ a_n=h(b_n) \end{cases}$$
solve it we have: 
$$\begin{cases} b_n=- \frac{1}{log(n)} \\ a_n=- \frac{1}{{log^2(n)}} \end{cases}$$

## Question 1 (b)
When $\xi < 0$ the GEV distribution has light upper tail with the finite upper limit which is  $\mu - \sigma / \xi$.

Let $\mu - \sigma / \xi = 0$ then this distribution with a finite upper end point of 0  has an upper end point of infinity.

## Question 1 (c)
We assume $A = \{ x: 0< F(x)<1 \}$ , and $x^*=sup_{x \in A} A$.

Here F(x) can be any function including the function $(*)$.

For $\forall x, x<x^*$, we have $Pr(M_n \leq x)=F^n(x) \to 0, as \  n \to \infty$.

For $\forall x, x \geq x^*$, we have $Pr(M_n \leq x)=F^n(x) \to 1 , as \ n \to \infty$.

This is mean whatever x or F(x), $M_n = 0 or 1$ when $n \to \infty$.

The $M_n$ is Degenerate distribution, which is useless.

That is why we need fit a GEV model.

## Question 1 (d)
The log-likelihood:
$$\xi_n \approx h^{'}(x)|_x=u(n) \ where \ u(n)=F^{-1}(1-1/n)$$
Thus $$\xi= h^{'}(x) = -2/ln(365) = -0.3400$$

@@

```{r}
# Add R code here (and similarly elsewhere)
plot(wm$waterYear,wm$Hs,bty="l",pch=20,xlab="year",ylab="wave height (in metres)")
#(wm$Hs~wm$meanNAO) 
plot(wm$meanNAO,wm$Hs,bty="l",pch=20,xlab="wave height (in metres)",ylab="the mean of NAO")
abline(lm(wm$Hs~wm$meanNAO), col="red") # regression line (wm$Hs~wm$meanNAO) 
lines(lowess(wm$Hs~wm$meanNAO), col="blue") # lowess line (wm$Hs~wm$meanNAO)
#(wm$Hs~wm$maxNAO)
plot(wm$maxNAO,wm$Hs,bty="l",pch=20,xlab="wave height (in metres)",ylab="the max of NAO")
abline(lm(wm$Hs~wm$maxNAO), col="red") # regression line (wm$Hs~wm$maxNAO)
lines(lowess(wm$Hs~wm$maxNAO), col="blue") # lowess line (wm$Hs~wm$maxNAO)
#(wm$Hs~wm$meanA0)
plot(wm$meanA0,wm$Hs,bty="l",pch=20,xlab="wave height (in metres)",ylab="the mean of AO")
abline(lm(wm$Hs~wm$meanA0), col="red") # regression line (wm$Hs~wm$meanA0)
lines(lowess(wm$Hs~wm$meanA0), col="blue") # lowess line (wm$Hs~wm$meanA0)
#(wm$Hs~wm$maxAO)
plot(wm$maxAO,wm$Hs,bty="l",pch=20,xlab="wave height (in metres)",ylab="the max of AO")
abline(lm(wm$Hs~wm$maxAO), col="red") # regression line (wm$Hs~wm$maxAO)
lines(lowess(wm$Hs~wm$maxAO), col="blue") # lowess line (wm$Hs~wm$maxAO)

```

## Storm peaks (pot)
```{r}
plot(pot$date,pot$Hs,bty="l",pch=20,xlab="date",ylab="wave height (in metres)")
```

## Comments
The relationship between wave height and all of variables can be roughly seen through the scatter plot. As can be seen from the line graph, the more the red line coincides with the blue line, the better the linear regression model fits. As can be seen from the plot, the average of AO has a better fitting relationship with the height of the waves. Further analysis needs to be carried out, followed by analysis.
${}$\hspace*{\fill} [10]

@@

## GEV modelling of winter maxima
```{r}
wm.gev <- gev.fit(wm[,1])#fit GEV model
wm.gev$mle#GEV MLEs
wm.gev$se#standard errors of MLEs
pjn.gev.diag(wm.gev) #PJN's GEV diagnostics
```

### Maximum Likelihood-Based Inference
```{r}
pjn.gev.conf(wm.gev, conf = 0.95)
pjn.gev.profxi(wm.gev, -1, 1, conf = 0.95)   
pjn.gev.conf.ret.levels(wm.gev, m = 100, conf = 0.95)
# Profile log-likelihood for the 100 year return level
pjn.gev.prof(wm.gev,m=100,14,20)# PJN's version
```

@@

First, the model is fitted, and the three parameters of the model are 11.2726903,  1.2064985, -0.1534719 respectively. Their variances are 0.18140907, 0.12880447, 0.09956538 respectively. Through the diagnosis, we can see that the tail of data is basically appropriate, and it can be seen the model's trend.

### Bayesian Inference
```{r}
my.ylab <- c("mu","sigma","xi")
# Set independent normal priors with large variances 
prior.cov.mat <- diag(c(10000, 10000, 100))         
pn <- prior.norm(mean = c(0, 0, 0), cov = prior.cov.mat)

# Starting well away from the MLEs
init <- c(11,1,0.1)                
# SDs of the proposal distributions for (mu, ln(sigma), xi)
prop.sd <- c(0.18,0.12,0.1)       
prop.sd.auto <- ar.choice(init = init, prior = pn, lh = "gev", 
                          data = wm$Hs, psd = prop.sd, 
                          tol = rep(0.02, 3))$psd
post <- posterior(3000, init = init, prior = pn, lh = "gev", 
                  data = wm$Hs, psd = prop.sd.auto)
for (i in 1:3) plot(post[,i],ylab=my.ylab[i]) 
attr(post,"ar")        
# Posterior dependence among parameters
pairs(post)
```

### Comments
Bayesian Inference

## Binomial-GP modelling of storm peaks
```{r}
pot.gp <- gpd.fit(pot[,1], 10)
```

### Threshold selection
```{r}
mrl.plot(pot[,1])
# PJN's GP diagnostics 
pjn.gpd.diag(pot.gp)    
```

### Maximum Likelihood-Based Inference
```{r}
#Symmetric 100*conf% confidence intervals for p_u, sigma_u and xi 
pjn.gpd.conf(pot.gp, conf = 0.95)
#PJN's version of ismev::gpd.profxi 
pjn.gpd.profxi(pot.gp, xlow = -0.3, xup = 0.2, conf = 0.95) 
#Symmetric 100*conf% confidence intervals for the 100-year return level 
pjn.gpd.conf.ret.levels(pot.gp, m = 100, npy = 10, conf = 0.95)
#Profile log-likelihood for the 100-year return level 
pjn.gpd.prof(pot.gp, m = 100, xlow = 14, xup = 20, npy = 10, conf = 0.95) 
```

@@

The estimated of $\xi$ is -0.15, and the corresponding 95% confidence interval is [-0.33,0.07].

### Bayesian Inference
```{r}
fp <- set_prior(prior = "flat", model = "gp", min_xi = -1)
bp <- set_bin_prior(prior = "jeffreys")
u <- quantile(pot[,1], probs = 0.8)
bgpg <- rpost(n = 10000, model = "bingp", prior = fp, thresh = u, data = pot[,1],
              bin_prior = bp, npy = 10)
plot(bgpg, pu_only = TRUE)
plot(bgpg, add_pu = FALSE)
plot(bgpg, add_pu = TRUE)

```

@@

It can also be seen that the likelihood function of the 100 year return level. It takes a constant attempt to return the plot to find th likelihood function is asymmetrical, the confidence interval is also asymmetrical with respect to the maximum likelihood estimate. 
${}$\hspace*{\fill} [25]

@@

From the plot we can see that the return period is 100. The estimated level of reconstruction for the year is 15.25, and the corresponding 95% confidence interval is [14.42,17.7].

@@

## Build a GEV regression model

${}$\hspace*{\fill} [15]
```{r}

evd_fit0 <- fgev(wm[,1])
covaryear <- data.frame(scaled_year = (wm[,2]-1955)/(2010-1955))
covar2 <- data.frame(scaled_year = (wm[,2]-1955)/(2010-1955), NOA = wm[, 3])
evd_fit1 <- fgev(wm[,1], nsloc = covaryear)
evd_fit2 <- fgev(wm[,1], nsloc = covar2)

anova(evd_fit2, evd_fit1, evd_fit0)

fitted(evd_fit2) # strictly speaking this should be coef() 
std.errors(evd_fit2)
vcov(evd_fit2)
confint(evd_fit2)
plot(evd_fit2)
```
